---
title: "gtsummary in quarto"
format: 
 html:
  theme: yeti
execute:
  echo: false
---

@tbl-one is a demographics table and @tbl-two is a distribution table


```{r}
#| message: false
#| echo: false
library(tidyverse)
library(gtsummary)
nlsy_cols <- c("glasses", "eyesight", "sleep_wkdy", "sleep_wknd", 
               "id", "nsibs", "samp", "race_eth", "sex", "region", 
               "income", "res_1980", "res_2002", "age_bir")
nlsy <- read_csv(here::here("data", "raw", "nlsy.csv"), 
                 na = c("-1", "-2", "-3", "-4", "-5", "-998"),
                 skip = 1, col_names = nlsy_cols) |> 
  mutate(region_cat = factor(region, labels = c("Northeast", "North Central", "South", "West")),
         sex_cat = factor(sex, labels = c("Male", "Female")),
         race_eth_cat = factor(race_eth, labels = c("Hispanic", "Black", "Non-Black, Non-Hispanic")),
         eyesight_cat = factor(eyesight, labels = c("Excellent", "Very good", "Good", "Fair", "Poor")),
         glasses_cat = factor(glasses, labels = c("No", "Yes"))) 
```

# Customization of `tbl_summary()`

```{r}
#| label: tbl-one
#| tbl-cap: "This is a demographics table"
tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir))
```

```{r}
#| label: tbl-two
#| tbl-cap: "This is a distribution table"
tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir),
  label = list(
    race_eth_cat ~ "Race/ethnicity",
    region_cat ~ "Region",
    eyesight_cat ~ "Eyesight",
    glasses ~ "Wears glasses",
    age_bir ~ "Age at first birth"
  ),
  missing_text = "Missing")
```

#customisation of `tbl_summary()`
```{r}
#| label: tbl-sex
#| tbl-cap: "Descriptive stats stratified by sex"
tbl_summary(
 nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir))
```

```{r}
#| label: tbl-eyesight
#| tbl-cap: "Descriptive stats stratified by eye sight"
tbl_summary(
 nlsy,
  by = eyesight_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir))
```


The descriptive stats are stratified by sex (@tbl-sex)and by eyesight(@tbl-eyesight).

#Add a chunk where you calculate the mean number of hours of sleep on weekends (you can use the mean() function). Have the code print that output.

```{r}
#| echo: true
mean_sleep_weekend <- mean(nlsy$sleep_wknd, na.rm = TRUE)
mean_sleep_weekend
```

The mean sleep hours on the weekend of the participants is `r mean_sleep_weekend`

```{r}
#| echo: true
median_sleep_weekend <- median(nlsy$sleep_wknd, na.rm = TRUE)
median_sleep_weekend
```
The mean sleep hours on the weekend of the participants is `r median_sleep_weekend`


The mean sleep hours on the weekend of the participants is `r round (mean_sleep_weekend, 2)` and the median was `r median_sleep_weekend`


```{r}
#| label: tbl-descr
#| tbl-cap: "Descriptive statistics"
table1 <- tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir)) |> 
  add_overall(last = TRUE)
table1
```
```{r}
inline_text(table1, variable = "age_bir", column = "stat_1")
```

```{r}
inline_text(table1, variable = "glasses", column = "stat_0",
						pattern = "{n}/{N} ({p}%)")
```

```{r}
#| label: tbl-income
#| tbl-cap: "Descriptive statistics"
income_table <- tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(income
 )) |> 
  add_overall(last = TRUE)
income_table
```
```{r}
inline_text(income_table, variable = "income", column = "stat_2")
```



```{r}
#| label: tbl-sex-sleep
#| tbl-cap: "Descriptive statistics"

nlsy <- filter (nlsy, glasses == 1)
table_sex_sleep <- tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sleep_wknd, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir
  						)) |> 
  add_overall(last = TRUE)
table_sex_sleep
excellent_eyesight <- inline_text(table_sex_sleep, variable = "eyesight_cat", level = "Excellent", column = "stat_0")
```

`r excellent_eyesight`people had excellent eyesight.
